Landcover Data Limit Our Understanding of Global Change

### 2015-12-01 paper/main.tex
Commit multiple changes to Appendix, including restructuring and renaming of Rmd files to faciliate creation of single code appendix, removal of many files from repo and into gitignore. Significance statement added to main ms.

\title{Landcover Data Limits Our Understanding of Global Change}
%\title{Our Understanding of Global Change is Built on a Shaky Foundation}

@@

%%%Newly updated.
%%% If significance statement need, then can use the below command otherwise just delete it.
%\significancetext{LDE developed the concept of the study, conducted the analysis, data interpretation and drafted the manuscript. \color{red}{To be completed}}

@@

The value of the insights resulting from such studies depends upon the veracity of their underlying landcover data, much as a house requires a solid foundation in order to remain standing. Unfortunately, the evidence to date indicates that much of our understanding of global change is built on shaky foundations. The reason for this is that landcover data can only practically be derived from satellite imaging, which has several important constraints--related to both physics and analytics--that propagate mapping errors. First, in many regions the cover types of interest are smaller than the sensor resolution \cite[e.g. smallholder's farms][]{jain_mapping_2013,debats_generalized_2015,ozdogan_resolution_2006}, or spectrally indistinct from neighboring covers \cite{sweeney_mapping_2015}. Second, the act of defining a cover class can cause error, in that selected classes may have highly diverse spectral properties (e.g. agriculture \cite{debats_generalized_2015,estes_platform_2015}) and thus be difficult for the classifier to distinguish, while discretizing a continuous cover type (e.g. dividing a forest into different canopy cover classes) can also promote classification error, particularly near class boundaries \cite{foody_status_2002}, as well as confusion about the actual extent of the cover type \cite{sexton_conservation_2015}. Furthermore, class definitions often vary between maps, complicating inter-comparison \cite{kuemmerle_challenges_2013}. Third, landcover map series are often used to detect changes \cite[e.g.][]{gross_monitoring_2013}, or may use within-class seasonality as a means for discriminating between classes (e.g. phenology of cropping versus surrounding vegetation \cite{sweeney_mapping_2015}), but seasonal variability and landcover changes can be easily confused. Given these multiple sources of error, landcover maps are often inaccurate at finer scales and disagree widely between products, particularly in the world's most rapidly developing regions \cite{estes_projected_2013, fritz_comparison_2010, fritz_cropland_2011}. These errors limit our ability to obtain granular, mechanistic understanding of global change processes.

These problems with landcover products are well known \cite{fritz_comparison_2010, fritz_cropland_2011, see_improved_2015, fritz_mapping_2015,verburg_challenges_2011}, and there are a variety of map improvement efforts underway \cite[e.g.][]{fritz_geo-wiki:_2012, estes_platform_2015}. What remains an open question is exactly how much the maps researchers typically use deviate from actual landcover, and how this in turn impacts our understanding of global change processes. Fully answering this question depends on having comprehensive, spatially representative ground truth data, which are unavailable for most parts of the world, particularly over Africa and other developing regions \cite{see_improved_2015}. Our current understanding of map accuracy is therefore built primarily on bottom-up tests made with a relatively small number of ground truth points (relative to the total mapped area), or from top-down "sanity checks" made in comparison to aggregated survey data.  In terms of the former, there are well-developed, best-practice guidelines for selecting and using ground-truth samples to robustly quantify map error, but these are not commonly applied by the broader community of landcover map users (often due to lack of data), which makes it challenging to compare how error varies within and between maps \cite{stehman_global_2012, olofsson_making_2013,olofsson_good_2014,foody_status_2002}. The latter form of assessment are made at coarser administrative levels relative to reported statistics \cite[e.g.][]{fritz_comparison_2010}, which can also be suspect \cite{carletto_emperor_2013,fao_action_2013}.

%Top-down accuracy assessments are also made from  "sanity checks" made in comparison to aggregated survey data. 
%
%
%
%This allows us to quantify between-map discrepancies  \cite[e.g.][]{fritz_comparison_2010, kaptue_tchuente_comparison_2011}, or to understand map fidelity to country-level statistics \cite[e.g.][]{fritz_comparison_2010}, but offers little direction for how to arrive at a true number.
%
%
%
%In practice, this means that we have an incomplete understanding of map error, and 
%
%
%
%
%There are also well-defined guidelines for quantifying and accounting for map accuracy \cite{stehman_global_2012, olofsson_good_2014, olofsson_good_2014}. However, inadequate assessment of map error and their impacts on subsequent analyses remains quite common within the community of landcover map users \cite{olofsson_making_2013,olofsson_good_2014}. There is also an open question  
%
%or from top-down "sanity checks" made in comparison to aggregated survey data.

%This allows us to quantify between-map discrepancies  \cite[e.g.][]{fritz_comparison_2010, kaptue_tchuente_comparison_2011}, or to understand map fidelity to country-level statistics \cite[e.g.][]{fritz_comparison_2010}, but offers little direction for how to arrive at a true number.

@@

Fortunately, the recent, explosive growth in public and private initiatives to develop new Earth observing capabilities, which range from small drones\footnote{e.g. 3DRobotics, DJIA} to new high resolution satellite arrays\footnote{Planet Labs, Skybox} and better mapping methods \cite{fritz_geo-wiki:_2012,estes_projected_2013, debats_generalized_2015}, are finally providing the means to more comprehensively interrogate the accuracy and biases in the landcover products that have become commonplace in global change research--and which are often used to make policy decisions \cite{searchinger_high_2015}.

@@

We used this dataset as a reference layer for evaluating four landcover products representative of the type commonly used in Earth Systems research. The first was South Africa's own 30 m resolution 2009 National Landcover map (SA-LC)\cite{sanbi_national_2009}, which is typical of the higher-resolution, Landsat-based maps that are typically available only for individual countries \cite[e.g.][]{fry_completion_2009}. The second and third were respectively the 300 m GlobCover 2009 \cite{arino_global_2012} and 500 m resolution 2011 MODIS Landcover products, which are widely used global-scale products \cite[e.g.][]{gross_monitoring_2013, shackelford_conservation_2015}. The fourth dataset was the new 1 km GeoWiki hybrid-fusion cropland map for Africa \cite{fritz_mapping_2015}, which incorporates the first three datasets and represents the current state-of-the-art in landcover mapping.  Since GeoWiki provides a continuous bounded value of landcover (percent cropland cover), which cannot be feasibly converted to a categorical value, we converted all other datasets, including the reference vectors, into comparable 1 km gridded percent cropland estimates.

@@

We then used these maps to conduct four further analyses typical of Earth System science research. The first two related to physical processes, namely the calculation of vegetative carbon stocks and the simulation of evapotranspiration by a hydrological model. The second two were socio-economic in nature: the estimation of agriculture yield and production, and an agent-based model assessment of household food security.  The first and third analyses were relatively simple, in that the variable(s) of interest were mapped onto landcover using empirical relationships. The second and fourth relied on more complex numerical methods, where landcover was one of several variables needed to run each model. For the simpler analyses, we examined how results were influenced by map aggregation, while for the more complex cases, our assessments were confined to each numerical model's standard output resolution.

@@

%\begin{figure*}[!b]
%\centering
%  \begin{minipage}[b]{.45\linewidth}
%    \centerline{\includegraphics[scale=0.51]{figures/carbon_veg_scalew.pdf}}
%    \caption{Caption}
%    \label{afoto3}
%  \end{minipage}\hfill%\qquad
%  \begin{minipage}[b]{.45\linewidth}
%    \centerline{\includegraphics[scale=0.51]{figures/et_bias_map.pdf}}
%    \caption{Caption 2}
%    \label{afoto4}
%  \end{minipage}
%\end{figure*}

%\vspace{-0.75 cm}
\begin{figure}[!h]
\centerline{\includegraphics[width=.5\textwidth]{figures/et_bias_map.pdf}}
\caption{Differences in annual mean evapotranspiration estimates from 29-year runs of the VIC land surface hydrology model when initialized with LAI response curves derived from the reference map, versus those from the four test maps.}\label{afoto}
\end{figure}

@@

\subsection{Agent-based simulation of food security}
Spatially-explicit agent-based model (ABMs) are frequently employed to understand land use decision-making and socio-ecological systems, often to facilitate improved policy, particularly in the arena of human development \cite{berger_creating_2006}. To obtain valid insight into the functioning of such systems, it is important to calibrate an ABM to empirical data describing the characteristics of land and land users, so that the model realistic represents the social and biophysical features of the study region \cite{berger_creating_2006}. In our example, we used an ABM of household food security that simulates food production by individual farming households (the agents) in response to weather and management capacity, which varies as a function of household resources, including the area and physical characteristics of fields \cite{chen_dependency_2013}. The model is initialized such that each household is allocated its designated share of cropland, and then produces a crop over the course of one to many seasons. Like many spatial ABMs, the model is computationally intensive, and thus run over smaller geographic domains (e.g. districts, rather than an entire country) and at higher spatial resolutions (10s to 100s of meters) in order to represent the different land units of single farmers. To match these computational characteristics, we selected four contiguous magisterial districts (ranging from 1,040-1,343 km$^2$, SI) in the eastern part of the country with similar climate and 28-45\% of their areas devoted to cropland. The initialization process iteratively assigns households to the landscape using a function that factors in neighbor and cropland proximity, to ensure that households are grouped into communities and that their fields are within a realistic proximity. The number of households and the cropland area per household was derived from surveys of farming communities where all cropland is used. The model was thus considered to be well-calibrated when all households were allocated their appropriate area of cropland, and all cropland was allocated to households.

@@

We used the reference map and each cropland map to separately initialize the model, and examined how map errors impacted the land allocation process and household food production estimates. We examined three variables: the percent of unallocated cropland, land deficit, and food deficit. The percent of unallocated cropland measures how effective the model was in matching household agents to available cropland resources. Where error was negative (indicating a cropland overestimate by the test maps), the amount of land left unallocated had a straight one-to-one relationship with the percentage of overestimation (Fig. 6A). When cropland was underestimated, no land was left unallocated until the underestimation exceeded 50\%. The MODIS-based simulation for districts 1 and 2 was most pronounced for this tendency, where 5-10\% of cropland remaining unallocated despite the fact that cropland with such a high cropland underestimate (85\%) the majority of households were not allocated cropland. This non-linear relationship occurred because croplands tend to cluster, and when underestimated clusters tend to be small, resulting in islands of cropland that fall outside of the search radius within which cropland is sought when agents are seeded onto the landscape.

The second measure, land deficit, describes the amount of land that the model should have assigned but couldn't due to the mismatch between the cropland map and the statistical estimate of cropland holdings. Land deficit increased exponentially in relation to cropland underestimation--reaching around 800\% for MODIS in districts 1 and 2 (Fig. 6B)--and would become infinite in the case of a 100\% underestimate. This contrasted with the third measure, food deficit, the shortfall in the average amount of food that should have been produced by each household but wasn't because of the land deficit; this particular deficit increased linearly with the percentage of cropland underestimate (Fig. 6C).

@@

\begin{figure}[!ht]
\centerline{\includegraphics[width=.5\textwidth]{figures/figure6.pdf}}
\caption{Biases in agent-based model results relative to the district-wise errors (as a percent) in total cropland area, in terms of A) the percent of cropland in each district that was not allocated to any household, B) the land deficit, or the percent of cropland that should have been allocated to households but wasn't, and C) the food deficit, or the average food production across households resulting from inadequate cropland allocation. Dot sizes correspond to district numbers, colors represent the landcover map.}
\label{afoto}
\end{figure}

@@

\indent The agent-based model of food security used here calculates average household-level food production over multiple seasons based on the biophysical environment (soil properties, weather), management inputs (planting decisions, labor availability), and farm assets (total field area) \cite{chen_dependency_2013}. The frequency distributions of household cropland holdings were derived from household survey data obtained from Zambia (no equivalent data were available for South Africa). We used these statistics to calculate the ``true'' number of households per district by dividing reference cropland areas by the mean cropland area (2.2 ha), and preserved the cropland area distributions by multiplying the total number of households by the frequencies. To create cropland surfaces for the, we disaggregated all five cropland maps to binary cropland/non-cropland cover types with 100 m resolution, in order to match the survey-recorded average field size of 1 ha. In initializing, the model takes a weighted (by cropland area frequency) random draw of 100 households and places these within a district, assigning each household its required number of ``fields'' (cropland pixels), which must be within 1.5 km of the household's location and not already assigned to another household. This process is iterated until all households are assigned cropland, or all available cropland is allocated. The model then runs separately for each district, with the household agents planting and managing a maize crop throughout the growing season, with possible final maize yields drawn from a look-up table that relates yield to particular climatic outcomes and management decisions. We ran the model for a single season, once per district and when initialized by each of the cropland maps (20 simulations total).  \\

@@

%The first of these was the widely used International Panel on Climate Change's Tier-1 approach for mapping vegetative carbon stocks, as developed by \cite{ruesch_new_2008}. The second was maize yield maps derived by disaggregating district-scale agricultural census data for both maize yield and harvested area \cite[following][]{monfreda_farming_2008,ramankutty_farming_2008}, from which we calculated the third map, gridded maize production estimates. Maps based on these analyses underpin many assessments of crop productivity and production \cite[e.g.][]{foley_solutions_2011,licker_mind_2010}.

%The reference dataset covered all of South Africa's field crop estate, from which we filtered out areas of communally owned farmlands and permanent tree crops to have a common basis of comparison across all landcover products, leaving us with map, after conversion to cropland percentage, that covered or 90\% of South Africa (1,081,000 km$^2$), of which 104,304 km$^2$ 2011 reference map showed a cropland area of 104,304 km$^2$, which the SA-LC and GeoWiki maps overestimated by 31 and 10\%, respectively, and GlobCover and MODIS underestimated by 18 and 23\%.

%The difference between total carbon stocks for the country made using any of the cropland maps were within +/-3\% of those based on the reference map, regardless of which cover type was adjacent to cropland (Table S1), as the large uncropped area of South Africa (Fig. 1) dilutes the errors within the 
%(which is constrained by an absolute distance and the proximity of other agents)

%\section{Digital RCD Analysis} 
\end{materials}

@@

% bib solution from here
% http://tex.stackexchange.com/questions/167650/is-there-a-more-recent-bibliography-style-file-bst-for-pnas
% https://github.com/jburon/pnas2011.bst
\bibliographystyle{pnas2011} 
{\footnotesize \bibliography{/Users/lestes/Dropbox/publications/full.bib}}

%\begin{thebibliography}{10}
%\bibitem{BN}
%M.~Belkin and P.~Niyogi, {\em Using manifold structure for partially
%  labelled classification}, Advances in NIPS, 15 (2003).
%
%\bibitem{BBG:EmbeddingRiemannianManifoldHeatKernel}
%P.~B\'erard, G.~Besson, and S.~Gallot, {\em Embedding {R}iemannian
%  manifolds by their heat kernel}, Geom. and Fun. Anal., 4 (1994),
%  pp.~374--398.
%\end{thebibliography}

>>>>

\title{Landcover Data Limit Our Understanding of Global Change}
%\title{Our Understanding of Global Change is Built on a Shaky Foundation}

@@

%%%Newly updated.
%%% If significance statement need, then can use the below command otherwise just delete it.
\significancetext{Remote sensing-derived landcover maps underpin much research into global environmental change, particularly in the most rapidly developing regions. These maps often contain substantial error, yet inadequate validation data makes it impossible to fully quantify map errors and how they affect our understanding of global change processes. Using a unique, high quality reference map of cropland distribution, this study comprehensively assesses landcover map bias and accuracy in a major agricultural region of sub-Saharan Africa, finding that map errors are strongly correlated with the landcover density, and can propagate substantial errors in downstream global change analyses. Users of landcover maps should select later generation products and aggregate to appropriate resolutions to avoid misleading insights and policies related to global change. }

@@

The value of the insights resulting from such studies depends upon the veracity of their underlying landcover data, much as a house requires a solid foundation in order to remain standing. Unfortunately, the evidence to date indicates that much of our understanding of global change is built on shaky foundations. The reason for this is that landcover data can only practically be derived from satellite imaging, which has several important constraints--related to both physics and analytics--that propagate mapping errors. First, in many regions the cover types of interest are smaller than the sensor resolution \cite[e.g. smallholder's farms][]{jain_mapping_2013,debats_generalized_2015,ozdogan_resolution_2006}, or spectrally indistinct from neighboring covers \cite{sweeney_mapping_2015}, which are factors that increase mapping complexity \cite{yu_meta-discoveries_2014}. Second, the act of defining a cover class can cause error, in that selected classes may have highly diverse spectral properties (e.g. agriculture \cite{debats_generalized_2015,estes_platform_2015}) and thus be difficult for the classifier to distinguish, while discretizing a continuous cover type (e.g. dividing a forest into different canopy cover classes) can also promote classification error, particularly near class boundaries \cite{foody_status_2002}, as well as confusion about the actual extent of the cover type \cite{sexton_conservation_2015}. Furthermore, class definitions often vary between maps, complicating inter-comparison \cite{kuemmerle_challenges_2013}. Third, landcover map series are often used to detect changes \cite[e.g.][]{gross_monitoring_2013}, or may use within-class seasonality as a means for discriminating between classes (e.g. phenology of cropping versus surrounding vegetation \cite{sweeney_mapping_2015}), but seasonal variability and landcover changes can be easily confused. Given these multiple sources of error, landcover maps are often inaccurate at finer scales and disagree widely between products, particularly in the world's most rapidly developing regions \cite{estes_projected_2013, fritz_comparison_2010, fritz_cropland_2011}. These errors limit our ability to obtain granular, mechanistic understanding of global change processes.

These problems with landcover products is known \cite{fritz_comparison_2010, fritz_cropland_2011, see_improved_2015, fritz_mapping_2015,verburg_challenges_2011}, and there are a variety of map improvement efforts underway \cite[e.g.][]{fritz_geo-wiki:_2012, estes_platform_2015}. What remains an open question is exactly how much the maps researchers typically use deviate from actual landcover, and how this in turn impacts our understanding of global change processes. Fully answering this question depends on having comprehensive, spatially representative ground truth data, which are typically unavailable for Africa and other rapidly changing regions \cite{see_improved_2015}, which have also been fairly sparsely mapped \cite{yu_meta-discoveries_2014}. Our current understanding of map accuracy over such areas is therefore based on a small number of maps \cite{yu_meta-discoveries_2014}, which were assessed using either bottom-up tests made with a relatively small number of ground truth points, or from top-down "sanity checks" made in comparison to aggregated survey data.  In terms of the former, there are well-developed, best-practice guidelines for selecting and using ground-truth samples to robustly quantify map error, but these are not commonly applied by the broader community of landcover map users (often due to lack of data), which makes it challenging to compare how error varies within and between maps \cite{stehman_global_2012, olofsson_making_2013,olofsson_good_2014,foody_status_2002}. The latter form of assessment are made at coarser administrative levels relative to reported statistics \cite[e.g.][]{fritz_comparison_2010}, which can also be suspect \cite{carletto_emperor_2013,fao_action_2013}.

@@

Fortunately, the recent, explosive growth in public and private initiatives to develop new Earth observing capabilities, which range from small drones\footnote{e.g. 3DRobotics, DJIA} to new high resolution satellite arrays\footnote{Planet Labs, Skybox} and better mapping methods \cite{fritz_geo-wiki:_2012,tuanmu_global_2014,estes_projected_2013, debats_generalized_2015}, are finally providing the means to more comprehensively interrogate the accuracy and biases in the landcover products that have become commonplace in global change research--and which are often used to make policy decisions \cite{searchinger_high_2015}.

@@

We used this dataset as a reference layer for evaluating four landcover products representative of the type commonly used in global change studies. The first was South Africa's own 30 m resolution 2009 National Landcover map (SA-LC)\cite{sanbi_national_2009}, which is typical of the higher-resolution, Landsat-based maps that are typically available only for individual countries \cite[e.g.][]{fry_completion_2009}. The second and third were respectively the 300 m GlobCover 2009 \cite{arino_global_2012} and 500 m resolution 2011 MODIS Landcover products, which are widely used global-scale products \cite[e.g.][]{gross_monitoring_2013, shackelford_conservation_2015}. The fourth dataset was the new 1 km GeoWiki hybrid-fusion cropland map for Africa \cite{fritz_mapping_2015}, which incorporates the first three datasets and represents the current state-of-the-art in landcover mapping.  Since GeoWiki provides a continuous bounded value of landcover (percent cropland cover), which cannot be feasibly converted to a categorical value, we converted all other datasets, including the reference vectors, into comparable 1 km gridded percent cropland estimates.

@@

We then used these maps to conduct four further analyses typical of global change research. The first two related to physical processes, namely the calculation of vegetative carbon stocks and the simulation of evapotranspiration. The second two were socio-economic in nature: the estimation of agriculture yield and production, and an agent-based model simulating household dynamics in dryland agroecosystems.  The first and third analyses were relatively simple, in that the variable(s) of interest were mapped onto landcover using empirical relationships. The second and fourth relied on more complex numerical methods, where landcover was one of several variables needed to run each model. For the simpler analyses, we examined how results were influenced by map aggregation, while for the more complex cases, our assessments were confined to each numerical model's standard output resolution.

@@

%\vspace{-0.75 cm}
\begin{figure}[h]
\centerline{\includegraphics[width=.5\textwidth]{figures/et_bias_map.pdf}}
\caption{Differences in annual mean evapotranspiration estimates from 29-year runs of the VIC land surface hydrology model when initialized with LAI response curves derived from the reference map, versus those from the four test maps.}\label{afoto}
\end{figure}

@@

\subsection{Agent-based simulation of households in dryland agroecosystems}
Spatially-explicit agent-based model (ABMs) are frequently employed to understand land use decision-making and socio-ecological systems, often to facilitate improved policy, particularly in the arena of human development \cite{berger_creating_2006}. To obtain valid insight into the functioning of such systems, it is important to calibrate an ABM to empirical data describing the characteristics of land and land users, so that the model realistic represents the social and biophysical features of the study region \cite{berger_creating_2006}. In our example, we used an ABM of household food security that simulates food production by individual farming households (the agents) in response to weather and management capacity, which varies as a function of household resources, including the area and physical characteristics of fields \cite{chen_dependency_2013}. The model is initialized such that each household is allocated its designated share of cropland based on aggregated farm-level statistics, and then calculates household-level agricultural production based on maize yields simulated by the DSSAT cropping system model \cite{jones_modelling_2003}. Yield estimates vary as a function of soil properties, management inputs, and weather. The former two factors differ between households, as do the total areas of cropland, thus total production can vary substantially between households. Like many spatial ABMs, the model is computationally intensive, and thus run over smaller geographic domains (e.g. districts, rather than an entire country) and at higher spatial resolutions (10s to 100s of meters) in order to represent the different land units of single farmers. To match these computational characteristics, we selected four contiguous magisterial districts (ranging from 1,040-1,343 km$^2$, SI) in the eastern part of the country with similar climate and 28-45\% of their areas devoted to cropland. The initialization process iteratively assigns households to the landscape using a function that factors in neighbor and cropland proximity, to ensure that households are grouped into communities and that their fields are within a realistic proximity. The number of households and the cropland area per household was derived from surveys of farming communities where all cropland is used. The model was thus considered to be well-calibrated when all households were allocated their appropriate area of cropland, and all cropland was allocated to households.

@@

We used the reference map and each cropland map to separately initialize the model, and examined how map errors impacted the land allocation process and household food production estimates. We examined three variables: the percent of unallocated cropland, land deficit, and food deficit. The percent of unallocated cropland what share of the district's total cropland area was not assigned to any household, and is a measure of how effective the model was in matching household agents to available cropland resources. Where error was negative (indicating a cropland overestimate by the test maps), the amount of land left unallocated had a straight one-to-one relationship with the percentage of overestimation (Fig. 6A). When cropland was underestimated, no land was left unallocated until the underestimation exceeded 50\%. The MODIS-based simulation for districts 1 and 2 was most pronounced for this tendency, where 5-10\% of cropland remaining unallocated despite the fact that cropland with such a high cropland underestimate (85\%) the majority of households were not allocated cropland. This non-linear relationship occurred because croplands tend to cluster, and when underestimated clusters tend to be small, resulting in islands of cropland that fall outside of the search radius within which cropland is sought when agents are seeded onto the landscape.

The second measure, land deficit, is calculated from the total area of cropland that should have been allocated to households in each district, but wasn't, due to the mismatch between the cropland map and the statistical estimates of what total cropland holdings should have been (given the number of households and the size class distributions of cropland in the district). Land deficit, expressed as a percentage of the total area of cropland in the district (thus values exceeded 100\% when total deficit exceeded cropland area),  increased exponentially in relation to cropland underestimation--reaching around 800\% for MODIS in districts 1 and 2 (Fig. 6B)--and would become infinite in the case of a 100\% underestimate. This contrasted with the third measure, food deficit, the percentage shortfall in the average amount of food production that should have been produced by each household but wasn't because of the land deficit; this particular deficit increased linearly with the percentage of cropland underestimate (Fig. 6C).

@@

\begin{figure}[!ht]
\centerline{\includegraphics[width=.5\textwidth]{figures/figure6.pdf}}
\caption{Biases in agent-based model results relative to the district-wise errors (as a percent) in total cropland area, in terms of A) the percent of cropland in each district that was not allocated to any household, B) the land deficit, or the total area of cropland that should have been allocated to households in each district but wasn't (expressed as a percent of total district cropland, as determined by test maps), and C) the food deficit, or the percentage shortfall (relative to the reference simulation) in mean household food production resulting from inadequate cropland allocation. Dot sizes correspond to district numbers, colors represent the landcover map.}
\label{afoto}
\end{figure}

@@

\indent The agent-based model of food security used here calculates average household-level food production over multiple seasons based on the biophysical environment (soil properties, weather), management inputs (planting decisions, labor availability), and farm assets (total field area) \cite{chen_dependency_2013}. The frequency distributions of household cropland holdings were derived from household survey data obtained from Zambia (no equivalent data were available for South Africa). We used these statistics to calculate the ``true'' number of households per district by dividing reference cropland areas by the mean cropland area (2.2 ha), and preserved the cropland area distributions by multiplying the total number of households by the frequencies. To create cropland surfaces for each district, we disaggregated all five cropland maps to binary cropland/non-cropland cover types with 100 m resolution, in order to match the survey-recorded average field size of 1 ha. In initializing, the model takes a weighted (by cropland area frequency) random draw of 100 households and places these within a district, assigning each household its required number of ``fields'' (cropland pixels), which must be within 1.5 km of the household's location and not already assigned to another household. This process is iterated until all households are assigned cropland, or all available cropland is allocated. The model then runs separately for each district, with the household agents planting and managing a maize crop throughout the growing season, with possible final maize yields drawn from a look-up table that relates yield to particular climatic outcomes and management decisions, as simulated by the DSSAT model. We ran each ABM simulation for a single season, once per district and when initialized by each of the cropland maps (20 simulations total).\\

@@

\end{materials}

@@

\bibliographystyle{pnas2011} 
{\footnotesize \bibliography{/Users/lestes/Dropbox/publications/full.bib}}
